The Radio Sky at Meter Wavelengths: $m$-Mode Analysis Imaging with the Owens Valley Long Wavelength Array

### 2017-07-25 paper.tex
continue to update text and use aastex 6.1

\documentclass[twocolumn]{aastex6}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}

@@

\author{
    Michael W. Eastwood,
    Gregg Hallinan,
    Twilight Sparkle,
    Rainbow Dash
}

\begin{abstract}
    I present all-sky maps of the sky from the Owens Valley Long Wavelength Array.
    These maps are created from the application of $m$-mode analysis.
\end{abstract}

@@

Current 21 cm cosmology experiments can be broadly separated into two classes: global experiments
that are aiming to detect the spectral signature of the cosmologically redshifted 21 cm transition
after averaging over the entire sky, and power spectrum experiments that incorporate angular
information to attempt to measure the spatial power spectrum of cosmological 21 cm perturbations.
Ongoing global experiments include EDGES \citep{2010Natur.468..796B, 2017ApJ...835...49M}, LEDA
\todo{cite Price et al. 2017}, BIGHORNS \citep{2015PASA...32....4S}, SCI-HI
\citep{2014ApJ...782L...9V}, and SARAS 2 \citep{2017arXiv170306647S}.  Ongoing power spectrum
experiments include PAPER/HERA \citep{2015ApJ...809...61A, 2016arXiv160607473D}, LOFAR
\citep{2017ApJ...838...65P}, and the MWA \citep{2016ApJ...833..102B}.

@@

\begin{figure*}[ht]
    \plotone{maps/2017-04-20-rgb-map-smaller}
    \caption{three color map}
    \label{fig:rgb}
\end{figure*}

@@

@@

For a transit telescope, the visibilities $V^{ij}_\nu$ are a periodic function of sidereal
time\footnote{
    This is not strictly true. Ionospheric fluctuations and non-sidereal sources (such as the sun)
    will violate this assumption. This paper will, however, demonstrate that the impact on the final
    maps is mild.
}.
Therefore it is a very natural operation to compute the Fourier transform of the visibilities with
respect to sidereal time $\phi\in[0,2\pi)$.
\begin{equation}
    V^{ij}_{m,\nu} = \int_0^{2\pi} V^{ij}_\nu(\phi)\exp\bigg(-im\phi\bigg)\,\d\phi
\end{equation}
The output of this Fourier transform is the set of $m$-modes $V^{ij}_{m,\nu}$ where
$m=0,\,\pm1,\,\pm2,\,\ldots$ is the Fourier conjugate variable to the sidereal time. The $m$-mode
corresponding to $m=0$ is a simple average of the visibilities over sidereal time. Similarly $m=1$
corresponds to the component of the visibilities that varies over half-day timescales. Larger values
of $m$ correspond to components that vary on quicker timescales.

\todo{introduce the baseline transfer function because I will want to use it in later sections}

@@

\begin{figure}[t]
    \plotone{figures/lcurve/lcurve}
    \caption{
        Example l-curve (from spw04)
    }
    \label{fig:lcurve}
\end{figure}

@@

This trade-off can be analyzed quantitatively by trialing several values of $\varepsilon$ and
computing $\|\b v - \b B\b a\|^2$ and $\|\b a\|^2$ for each trial. An example is shown in
Figure~\ref{fig:lcurve}. The shape of this curve has a characteristic L-shape, and as a result this
type of plot is called an L-curve \todo{reference?}. The ideal value of $\varepsilon$ lies near the
turning point of the plot. At this point a small decrease in $\varepsilon$ will lead to an undesired
rapid increase in $\|\b a\|^2$, and a small increase in $\varepsilon$ will lead to an undesired
rapid increase in $\|\b v - \b B\b a\|^2$.

@@

While $m$-mode analysis imaging automatically deconvolves large angular scales, point sources do not
get deconvolved by the imaging process. This is a limitation of the set of basis functions we have
chosen to represent the sky -- the spherical harmonics. Point sources carry power to large values of
$l$, but the regularization will eventually begin to downweight spherical harmonics with $l$ larger
than what the interferometer is sensitive to. This consequence of this is that point sources will
appear in the maps with some characteristic point spread function. This point spread function may be
computed with
\begin{equation}
    \b a_\text{PSF}(\theta, \phi)
        = (\b B^*\b B + \varepsilon\b I)^{-1}\b B^*\b B\b a_\text{PS}(\theta, \phi)\,,
\end{equation}
where $\b a_\text{PSF}(\theta, \phi)$ is the vector of spherical harmonic coefficients representing
the point spread function at the spherical coordinates $(\theta, \phi)$, and $\b a_\text{PS}(\theta,
\phi)$ is the vector of spherical harmonic coefficients for a point source at $(\theta, \phi)$ given
by
\begin{equation}
    a_{lm, \text{PS}}(\theta, \phi) = Y_{lm}^*(\theta, \phi)\,.
\end{equation}

\todo{include a few images of the computed PSF for my maps}

\todo{describe how we incorporate this into major and minor iterations}

@@

The Owens Valley Long Wavelength Array (OVRO-LWA) is a 288-element interferometer located at the
Owens Valley Radio Observatory (OVRO) near Big Pine, California.  The OVRO-LWA is a low-frequency
instrument with instantaneous bandwidth covering 27.384 MHz to 84.912 MHz and 24 kHz channelization.
Each antenna stand hosts two perpendicular broadband dipoles so that there are $288\times2$ signal
paths in total. These signal paths feed into the 512-input LEDA correlator
\citep{2015JAI.....450003K}, which allows the OVRO-LWA to capture the entire visible hemisphere in a
single snapshot image.  In the current configuration 32 antennas (64 signal paths) are unused.

The 288 antennas are arranged in a pseudo-random configuration optimized to minimize sidelobes in
snapshot imaging.  251 of the antennas are contained within a 200 m diameter core. 32 antennas are
placed outside of the core in order to extend the maximum baseline length out to $\sim$1.5 km. The
final 5 antennas are equipped with noise-switched front ends for calibrated total power measurements
of the global sky brightness.  These antennas are used as part of the LEDA experiment \todo{cite
Price et al. 2017} to measure the global signal of 21 cm absorption from the cosmic dawn.
Figure~\ref{fig:antenna-layout} is a diagram of the antenna configuration.

\todo{cite Gregg's instrument paper}

@@

\todo{describe the ionospheric conditions at the time}

@@

@@

\section{A Nearby Supernova Remnant????}

@@

These maps are pretty cool. Any questions?

\todo{Cite Abhilash, Eastwood 2017}

>>>>

\documentclass[twocolumn]{aastex61}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage[]{algorithm2e}

@@

\author{Michael W. Eastwood}
\affiliation{Department of Astronomy, California Institute of Technology, 1200 E California Blvd, Pasadena, CA 91125}

\author{Marin M. Anderson}
\affiliation{Department of Astronomy, California Institute of Technology, 1200 E California Blvd, Pasadena, CA 91125}

\author{Ryan Monroe}
\affiliation{Department of Electrical Engineering, California Institute of Technology, 1200 E California Blvd, Pasadena, CA 91125}

\author{Gregg Hallinan}
\affiliation{Department of Astronomy, California Institute of Technology, 1200 E California Blvd, Pasadena, CA 91125}

\author{Esayas B. Shume} % contributed ionosphere text
\affiliation{Jet Propulsion Laboratory, 4800 Oak Grove Dr, Pasadena, CA 91109}

\begin{abstract}
    I present all-sky maps of the sky from the Owens Valley Long Wavelength Array.
    These maps are created from the application of $m$-mode analysis. \todo{Wednesday}
\end{abstract}

\keywords{
    cosmology: observations --
    dark ages, reionization, first stars --
    radio continuum: galaxies --
    radio continuum: ISM
}

@@

Current 21 cm cosmology experiments can be broadly separated into two classes: global experiments
that are aiming to detect the spectral signature of the cosmologically redshifted 21 cm transition
after averaging over the entire sky, and power spectrum experiments that incorporate angular
information to attempt to measure the spatial power spectrum of cosmological 21 cm perturbations.
Ongoing global experiments include EDGES \citep{2010Natur.468..796B, 2017ApJ...835...49M}, LEDA
\citep{todo_price_2017}, BIGHORNS \citep{2015PASA...32....4S}, SCI-HI \citep{2014ApJ...782L...9V},
and SARAS 2 \citep{2017arXiv170306647S}.  Ongoing power spectrum experiments include PAPER/HERA
\citep{2015ApJ...809...61A, 2016arXiv160607473D}, LOFAR \citep{2017ApJ...838...65P}, and the MWA
\citep{2016ApJ...833..102B}.

@@

\begin{figure*}[ht]
    \plotone{maps/2017-04-20-rgb-map-smaller}
    \caption{
        This Mollweide-projected map is constructed from 3 maps of the sky at 41.760 MHz (red),
        57.456 MHz (green), and 73.152 MHz (blue) using data from the Owens Valley Long Wavelength
        Array. These maps were constructed using $m$-mode analysis imaging. 8 bright radio sources
        have been subtracted from the sky (Cyg A, Cas A, Vir A, Tau A, Hya A, Her A, Per B, and 3C
        353). The hole corresponds to declinations less than $-30^\circ$. Most of the diffuse
        emission is synchrotron, but the blue regions mottling the galactic plane are due to
        free-free absorption from \ion{H}{2} regions.
    }
    \label{fig:rgb}
\end{figure*}

@@

For later convenience we will define the baseline transfer function $B^{ij}_\nu(\hat r)$ such that
\begin{equation}\label{eq:baseline-transfer-function}
    V^{ij}_\nu = \int_\text{sky} B^{ij}_\nu(\hat r) I_\nu(\hat r) \,\d\Omega \, .
\end{equation}
The baseline transfer function defines the response of a single baseline to the sky, and is a
function of the antenna primary beam and baseline orientation.

@@

For a transit telescope, the visibilities $V^{ij}_\nu$ are a periodic function of sidereal
time.\footnote{
    This is not strictly true. Ionospheric fluctuations and non-sidereal sources (such as the sun)
    will violate this assumption. This paper will, however, demonstrate that the impact on the final
    maps is mild.
}
Therefore it is a very natural operation to compute the Fourier transform of the visibilities with
respect to sidereal time $\phi\in[0,2\pi)$.
\begin{equation}
    V^{ij}_{m,\nu} = \int_0^{2\pi} V^{ij}_\nu(\phi)\exp\bigg(-im\phi\bigg)\,\d\phi
\end{equation}
The output of this Fourier transform is the set of $m$-modes $V^{ij}_{m,\nu}$ where
$m=0,\,\pm1,\,\pm2,\,\ldots$ is the Fourier conjugate variable to the sidereal time. The $m$-mode
corresponding to $m=0$ is a simple average of the visibilities over sidereal time. Similarly $m=1$
corresponds to the component of the visibilities that varies over half-day timescales. Larger values
of $m$ correspond to components that vary on quicker timescales.

@@

\begin{figure}[t]
    \plotone{figures/lcurve/lcurve}
    \caption{
        An example L-curve using OVRO-LWA data at 36.528 MHz. The $x$-axis is the norm of the
        solution (in this case the spherical harmonic coefficients) given in arbitrary units. The
        $y$-axis is the least-squares norm given in arbitrary units. $\varepsilon$ is the
        regularization parameter. When the regularization parameter is small, the norm of the
        solution grows rapidly. When the regularization parameter is large, the least-squares norm
        grows rapidly. The optimal value for the regularization parameter is near
        $\varepsilon=0.01$.
    }
    \label{fig:lcurve}
\end{figure}

@@

This trade-off can be analyzed quantitatively by trialing several values of $\varepsilon$ and
computing $\|\b v - \b B\b a\|^2$ and $\|\b a\|^2$ for each trial. An example is shown in
Figure~\ref{fig:lcurve}. The shape of this curve has a characteristic L-shape, and as a result this
type of plot is called an L-curve. The ideal value of $\varepsilon$ lies near the turning point of
the plot. At this point a small decrease in $\varepsilon$ will lead to an undesired rapid increase
in $\|\b a\|^2$, and a small increase in $\varepsilon$ will lead to an undesired rapid increase in
$\|\b v - \b B\b a\|^2$.

@@

$m$-mode analysis imaging automatically deconvolves angular scales described by modes for which the
singular values of $\b B$ are $\gg \sqrt{\varepsilon}$. For example, structures on angular scales
smaller than $\lambda/b_\text{min}$ and larger than $\lambda/b_\text{max}$ (where $b_\text{min}$ and
$b_\text{max}$ are the minimum and maximum baseline lengths) are generally well-measured by the
interferometer and will be automatically deconvolved during the imaging process. However, as the
total integration time of the data set increases, the noise in the data decreases. One consequence
of this is that the regularization parameter $\varepsilon$ can be decreased. This increases the
number of modes for which $\sigma_i \gg \sqrt{\varepsilon}$, and as a consequence the automatic
deconvolution can extend to scales larger than $\lambda/b_\text{min}$ and smaller than
$\lambda/b_\text{max}$.

However point sources do not get deconvolved by the imaging process even in the limit of infinite
SNR and $\varepsilon \rightarrow 0$. This is because point sources carry power to large values of
$l$ and by computational necessity we truncated the spherical harmonic representation of the sky at
some $l_\text{max}$. A consequence of this is that point sources will appear in the maps with a
characteristic point spread function (PSF). The PSF may be computed with:
\begin{equation}
    \b a_\text{PSF}(\theta, \phi)
        = (\b B^*\b B + \varepsilon\b I)^{-1}\b B^*\b B\b a_\text{PS}(\theta, \phi)\,,
\end{equation}
where $\b a_\text{PSF}(\theta, \phi)$ is the vector of spherical harmonic coefficients representing
the PSF at the spherical coordinates $(\theta, \phi)$, and $\b a_\text{PS}(\theta, \phi)$ is the
vector of spherical harmonic coefficients for a point source at $(\theta, \phi)$ given by
\begin{equation}
    a_{lm, \text{PS}}(\theta, \phi) = Y_{lm}^*(\theta, \phi)\,.
\end{equation}

In general the PSF can be a function of the right ascension and declination. However point sources
at the same declination take the same track through the sky and therefore (barring any ionospheric
effects) will have the same PSF. The PSF is actually only a function of the declination. For
example, sources at low elevations will tend to have an extended PSF along the north-south axis due
to baseline foreshortening. An example computed PSF is shown in Figure~\ref{fig:psf}.

The computed PSF as a function of frequency and declination can be used to adapt the CLEAN algorithm
\citep{1974A&AS...15..417H} to $m$-mode analysis.

\begin{algorithm}
    \KwData{dirty maps generated from the output of Equation~\ref{eq:tikhonov-solution};}
    \KwResult{clean maps with point sources deconvolved;}
    pre-compute $\b M = (\b B^*\b B + \varepsilon\b I)^{-1}\b B^*\b B$\;
    \While{noise in map $>$ threshold}{
        identify the pixel with the largest amplitude\;
        compute $\b a_\text{PSF} = \b M \b a_\text{PS}$\;
        update $\b a \leftarrow \b a - (\textit{gain})\times\b a_\text{PSF}$\;
    }
    convolve removed components with a CLEAN-beam and restore to the map\;
\end{algorithm}

It should be noted that for interferometers with more baselines than the number of spherical
harmonics used in the maps, $\b M$ is a much smaller matrix than the full transfer matrix $\b B$.
Therefore the cost of multiplying by $\b M$ can actually be dominated by the cost of computing the
spherical harmonics describing a point source at the identified pixel.

@@

\begin{figure}[t]
    \plotone{figures/psf/spw04-psf-45-degrees}
    \caption{
        The computed PSF for the OVRO-LWA for a point source at a declination of $+45^\circ$ and
        frequency of $36.528$ MHz.
    }
    \label{fig:psf}
\end{figure}

The Owens Valley Long Wavelength Array (OVRO-LWA) is a 288-element interferometer located at the
Owens Valley Radio Observatory (OVRO) near Big Pine, California \citep{todo_hallinan_2017}.  The
OVRO-LWA is a low-frequency instrument with instantaneous bandwidth covering 27.384 MHz to 84.912
MHz and 24 kHz channelization.  Each antenna stand hosts two perpendicular broadband dipoles so that
there are $288\times2$ signal paths in total. These signal paths feed into the 512-input LEDA
correlator \citep{2015JAI.....450003K}, which allows the OVRO-LWA to capture the entire visible
hemisphere in a single snapshot image.  In the current configuration 32 antennas (64 signal paths)
are unused.

The 288 antennas are arranged in a pseudo-random configuration optimized to minimize sidelobes in
snapshot imaging.  251 of the antennas are contained within a 200 m diameter core. 32 antennas are
placed outside of the core in order to extend the maximum baseline length out to $\sim$1.5 km. The
final 5 antennas are equipped with noise-switched front ends for calibrated total power measurements
of the global sky brightness.  These antennas are used as part of the LEDA experiment
\citep{todo_price_2017} to measure the global signal of 21 cm absorption from the cosmic dawn.
Figure~\ref{fig:antenna-layout} is a diagram of the antenna configuration.

@@

The geomagnetic conditions during this time were mild. The Disturbance storm time (Dst) index was
$>-30$ nT during the entirety of the observing period.\footnote{
    The Dst index was obtained from the World Data Center for Geomagnetism, Kyoto University
    (\url{http://swdcwww.kugi.kyoto-u.ac.jp/}).
}
Following the classification scheme of \citet{2008GMS...181.....K}, a weak geomagnetic storm has
$\text{Dst} < -30$ nT. Therefore there were no geomagnetic storms during the time of these
observations.

@@

\todo{tuesday}

@@

@@

\citet{todo_mishra_2017}

\todo{Wednesday}
